canaper: Categorical analysis of neo- and paleo-endemism in R

### 2022-09-21 ms/ms.Qmd
Update MS
format structure for MEE

Here, we describe a new R package that implements CANAPE completely in R, `canaper`.
We strived to make `canaper` simple to use and efficient.
Parallel computing can be enabled with a single line of code.
Randomization algorithms are provided via the `vegan` package [@Oksanen2017], which comes with \> 30 pre-defined algorithms as well as the option to provide a custom, user-defined algorithm.
`canaper` has undergone thorough code review through ROpenSci [@Boettiger2015] and is verified against a large number of unit tests (>98% coverage).
All results are reproducible by setting the random seed generator in R, in both sequential and parallel computing modes.

@@

Community data is provided as a `data.frame` or matrix, with species as columns and communities (also referred to as "sites" or "grid-cells") as rows.
In this case, the data must include both row names and column names.
Community data may also be input as a tibble, in which case site names must be indicated in a dedicated column (default column name `"site"`), rather than row names since tibbles lack row names.
Community data may be either presence-absence data (0s or 1s) or abundance data (integers \>= 0).
However, all calculations of PD and PE only use presence-absence information (i.e., no abundance weighting is used), so identical results will be obtained whether the input data is abundance or abundance that has been converted to presence-absence.
Community data is typically loaded using `read.csv()`, `readr::read_csv()`, or other functions that can import rectangular data.

@@

# Conclusion

@@

@@

# Figures
```{r}
#| label: fig-acacia-cpr
#| fig.height: 8
#| fig-cap: "Categorical analysis of neo- and paleo-endemism (CANAPE) of Australian *Acacia*."
#| echo: false

\(a) Map of Australia showing grid-cells (communities) colored by endemism type. (b) Scatterplot of comparing phylogenetic endemism (PE) of each community as measured on the original tree vs. a comparison tree with all non-zero branch lengths set to equal length, colored according to endemism type. This figure reproduces Figure 3 of @Mishler2014.

@@

@@

>>>>

Here, we describe a new R package that implements CANAPE completely in R, `canaper`.
We strived to make `canaper` simple to use and efficient.
Parallel computing can be enabled with a single line of code.
Randomization algorithms are provided via the `vegan` package [@Oksanen2017], which comes with \> 30 pre-defined algorithms as well as the option to provide a custom, user-defined algorithm.
`canaper` has undergone thorough code review through ROpenSci [@Boettiger2015] and is verified against a large number of unit tests (>95% coverage).
All results are reproducible by setting the random seed generator in R, in both sequential and parallel computing modes.

@@

Community data is provided as a `data.frame` or matrix, with species as columns and communities (also referred to as "sites" or "grid-cells") as rows.
In this case, the data must include both row names and column names.
Community data may also be input as a tibble, in which case site names must be indicated in a dedicated column (default column name `"site"`), rather than row names since tibbles lack row names.
Community data may be either presence-absence data (0s or 1s) or abundance data (integers \>= 0).
However, when PD and PE are calculated only presence-absence information is used (i.e., abundance weighting is not used), so identical results will be obtained whether the input data is abundance or abundance that has been converted to presence-absence.
Community data is typically loaded using `read.csv()`, `readr::read_csv()`, or other functions that can import rectangular data.

@@

# Conclusions

@@

# Acknowledgements

This research supported in part by Japan Society for the Promotion of Science (Kakenhi) Grant numbers 16H06279 and 22H04925.
Members of the Iwasaki Lab (The University of Tokyo) provided helpful comments while the package was in development.

# Conflict of interest

All authors declare no conflict of interest.

# Author contributions

Joel H. Nitta conceived the ideas and designed methodology; Joel H. Nitta and Shawn W. Laffan analysed the data; Joel H. Nitta led the writing of the manuscript.
All authors contributed critically to the drafts and gave final approval for publication.

# Data availability

All data and code are available in the `canaper` R package, which is distributed by the Comprehensive R Archive Network (CRAN).
Source code is available on GitHub (<https://github.com/ropensci/canaper>) and Zenodo (<https://doi.org/10.5281/zenodo.5094032>).
Usage of `canaper` is documented at https://docs.ropensci.org/canaper/.
The `acacia` dataset was originally published by @Mishler2014.

@@

{{< pagebreak >}}

# Figure captions

**@fig-acacia-cpr.** Categorical analysis of neo- and paleo-endemism (CANAPE) of Australian *Acacia*.
(a) Map of Australia showing grid-cells (communities) colored by endemism type.
(b) Scatterplot of comparing phylogenetic endemism (PE) of each community as measured on the original tree vs. a comparison tree with all non-zero branch lengths set to equal length, colored according to endemism type. This figure reproduces Figure 3 of @Mishler2014.

{{< pagebreak >}}

@@

{{< pagebreak >}}

@@

```{r}
#| label: fig-acacia-cpr
#| fig.height: 8
#| fig-cap: ""
#| echo: false
